Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2021-01-31 thesis/thesis.tex

@@

There are several clustering algorithms: KMeans, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
While each one behaves better according to the distribution of the objects to clusterize,
we have chosen KMeans by its simplicity and good results.

KMeans requires a single parameter, the number of clusters to build (N).
The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
KMeans starts by making an initial groups configuration and then,
it reassigns objects to other groups iteratively by minimizing the distance among points
inside the new group and maximizing the distance with the centers of the other groups.

@@

This algorithm works well and gives good results at first approach.
However, we need to set a large number of clusters to find the open cluster we are looking for.
This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
For that reason, we have search for a KMeans refinement based on an artificial neural network.

The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
or \emph{DEC} \cite{xie2016unsupervised} takes KMeans as its starting point,
but then, it trains an autoencoder to reduce the feature space
and pass this transformed data through a Clustering Layer which refines the previous selection.

@@

Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods:
SIMBAD Astronomical Database - CDS (Strasbourg),
Clusterix+TOPCAT method, and the results obtained for each step of this method (KMeans, DEC and DEC filtered).
Errors associated to KMeans and DEC predictions have been calculated as the \emph{Standard Error},
$SE = \sigma / \sqrt{N}$, where $N$ is the number of samples in the selected group.

\vfill
\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr}\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
        19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
      KMeans & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
      \textbf{DEC (filtered)} & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
    \end{tabular}
    \caption{Parameters shown are proper motion in right ascension and declination, parallax
             with their respective deviations and number of stars corresponding to Melotte 22 data.}
    \label{tab:results_melotte_22}
  \end{center}
\end{table}
\vfill

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & 19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
      \textbf{Clusterix+TOPCAT} & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
      KMeans & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
      DEC (filtered) & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
    \end{tabular}
  \end{center}
\end{table}

@@

Until now, we have focused our study in the open cluster Melotte 22.
Since our aim is to get a model that fits well for a wide range of clusters,
we will now proceed to show some results obtained for a selection of clusters with different topologies.

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/pm_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/parallax_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/hr_diagram_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 Clusterix+TOPCAT characterization.}
  \label{fig:result_ngc_2516_clusterix}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_pm_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 KMeans characterization. Identified as cluster \emph{g3}.}
  \label{fig:result_ngc_2516_kmeans}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 DEC characterization. Identified as cluster \emph{g3}.}
  \label{fig:result_ngc_2516_dec}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_filtered_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 DEC (filtered) characterization. Identified as cluster \emph{g3}.}
  \label{fig:result_ngc_2516_dec_filtered}
\end{figure}

@@

Figure \ref{fig:result_ngc_2516_clusterix} shows NGC 2516 characterized using Clusterix+TOPCAT tools.
Figure \ref{fig:result_ngc_2516_kmeans} shows eigth clusters identified by KMeans.
In this case, cluster \emph{g3} is the one we are looking for.
Figures \ref{fig:result_ngc_2516_dec} and \ref{fig:result_ngc_2516_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
Again, the cluster of interest is the group \emph{g3}.
Although in general, groups between KMeans model and DEC models may not match.

@@

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & -4.6579 $\pm$ 0.0075 & 11.1517 $\pm$ 0.0075 & 2.4118 $\pm$ 0.0006 & 1727 \\
      Clusterix+TOPCAT & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
      KMeans & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
      DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
      \textbf{DEC (filtered)} & -4.502 $\pm$ 0.14 & 10.114 $\pm$ 0.17 & 2.392 $\pm$ 0.004 & 1072 \\
    \end{tabular}
    \caption{NGC 2516 results.}
    \label{tab:results_ngc_2516}
  \end{center}
\end{table}

@@

\begin{figure}[H]
  \vfil
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 Clusterix+TOPCAT characterization.}
  \label{fig:result_ngc_2632_clusterix}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 KMeans characterization. Identified as cluster \emph{g1}.}
  \label{fig:result_ngc_2632_kmeans}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 DEC characterization. Identified as cluster \emph{g1}.}
  \label{fig:result_ngc_2632_dec}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 DEC (filtered) characterization. Identified as cluster \emph{g1}.}
  \label{fig:result_ngc_2632_dec_filtered}
\end{figure}

@@

Figure \ref{fig:result_ngc_2632_clusterix} shows NGC 2632 characterized by the validation method.
Figure \ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by KMeans.
Cluster \emph{g1} is the open cluster we are looking for.
Figures \ref{fig:result_ngc_2632_dec} and \ref{fig:result_ngc_2632_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
Open cluster NGC 2632 is labeled as group \emph{g1}.

@@

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & -36.047 $\pm$ 0.110 & -12.917 $\pm$ 0.066 & 5.371 $\pm$ 0.003 & - \\
      Clusterix+TOPCAT & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
      KMeans & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
      DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
      \textbf{DEC (filtered)} & -21.571 $\pm$ 0.74 & -14.234 $\pm$ 0.61 & 4.719 $\pm$ 0.03 & 714 \\
    \end{tabular}
    \caption{NGC 2632 results.}
    \label{tab:results_ngc_2632}
  \end{center}
\end{table}

@@

Figure \ref{fig:result_ngc_2682_clusterix} shows NGC 2682 characterized using Clusterix+TOPCAT tools.
Figure \ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by KMeans.
KMeans labels NGC 2682 as group \emph{g4}.
Figures \ref{fig:result_ngc_2682_dec} and \ref{fig:result_ngc_2682_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
The cluster of interest is the group \emph{g2}.

@@

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & -10.9737 $\pm$ 0.0064 & -2.9396 $\pm$ 0.0063 & 1.1325 $\pm$ 0.0011 & 1194 \\
      Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
      KMeans & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
      DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
      \textbf{DEC (filtered)} & -9.619 $\pm$ 0.13 & -3.317 $\pm$ 0.13 & 1.140 $\pm$ 0.003 & 990 \\
    \end{tabular}
    \caption{NGC 2682 results.}
    \label{tab:results_ngc_2682}
  \end{center}
\end{table}

@@

Figure \ref{fig:result_melotte_25_clusterix} shows Melotte 25 characterized using Clusterix+TOPCAT tools.
Figure \ref{fig:result_melotte_25_kmeans} shows nine clusters identified by KMeans.
KMeans labels Melotte 25 as group \emph{g3}.
Figures \ref{fig:result_melotte_25_dec} and \ref{fig:result_melotte_25_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
Melotte 25 is labeled as group \emph{g5} by DEC model.

@@

We have applied our proposed method to clusters with different topologies:

>>>>

This research has made use of the SIMBAD database, operated at CDS, Strasbourg, France
\cite[2000,A\&AS,143,9]{wenger2000simbad}, "The SIMBAD astronomical database", Wenger et al.

@@

There are several clustering algorithms: K-Means, Mean-Shift Clustering, DBSCAN \cite{ester1996density} among other.
While each one behaves better according to the distribution of the objects to clusterize,
we have chosen K-Means by its simplicity and good results.

K-Means requires a single parameter, the number of clusters to build (N).
The algorithm looks for N center points which are vectors of the same dimension as the number of selected features.
K-Means starts by making an initial groups configuration and then,
it reassigns objects to other groups iteratively by minimizing the distance among points
inside the new group and maximizing the distance with the centers of the other groups.

@@

This algorithm works well and gives good results at first approach.
However, we need to set a large number of clusters to find the open cluster we are looking for.
This fact complicates the identification of the OC and many times the found cluster contains to many outliers.
For that reason, we have search for a K-Means refinement based on an artificial neural network.

The \emph{Unsupervised Deep Embedding for Clustering Analysis} model
or \emph{DEC} \cite{xie2016unsupervised} takes K-Means as its starting point,
but then, it trains an autoencoder to reduce the feature space
and pass this transformed data through a Clustering Layer which refines the previous selection.

@@

Table \ref{tab:results_melotte_22} shows a summary of Melotte 22 with values taken from different sources and methods:
SIMBAD Astronomical Database - CDS (Strasbourg),
Clusterix+TOPCAT method, and the results obtained for each step of this method (K-Means, DEC and DEC filtered).
Errors associated to K-Means and DEC predictions have been calculated as the \emph{Standard Error},
$SE = \sigma / \sqrt{N}$, where $N$ is the number of samples in the selected group.

\vfill
\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr}\tablefootnote{Results have been taken from \protect\citeA[p.~25. Table A.3. Pleiades]{babusiaux2018gaia}} &
        19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
      Clusterix+TOPCAT & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
      K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
      \textbf{DEC (filtered)} & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
    \end{tabular}
    \caption{Parameters shown are proper motion in right ascension and declination, parallax
             with their respective deviations and number of stars corresponding to Melotte 22 data.}
    \label{tab:results_melotte_22}
  \end{center}
\end{table}
\vfill

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & 19.997 $\pm$ 0.127 & -45.548 $\pm$ 0.101 & 7.364 $\pm$ 0.005 & 1326 \\
      \textbf{Clusterix+TOPCAT} & 19.98 $\pm$ 1.25 & -45.47 $\pm$ 1.48 & 7.33 $\pm$ 0.21 & 634 \\
      K-Means & 20.25 $\pm$ 0.95 & -38.01 $\pm$ 1.08 & 7.23 $\pm$ 0.06 & 1378 \\
      DEC & 23.67 $\pm$ 1.29 & -46.23 $\pm$ 1.50 & 8.04 $\pm$ 0.09 & 878 \\
      DEC (filtered) & 19.50 $\pm$ 0.41 & -44.23 $\pm$ 0.39 & 7.42 $\pm$ 0.005 & 438 \\
    \end{tabular}
  \end{center}
\end{table}

@@

Until now, we have focused our study in the open cluster Melotte 22.
Since our aim is to get a model that fits well for a wide range of clusters,
we will now proceed to show some results obtained for a selection of clusters with different typologies.

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/pm_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/parallax_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/hr_diagram_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 Clusterix+TOPCAT characterization.}
  \label{fig:result_ngc_2516_clusterix}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_pm_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_parallax_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/kmeans_hr_diagram_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 K-Means characterization. Identified as cluster \emph{g3}.}
  \label{fig:result_ngc_2516_kmeans}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 DEC characterization. Identified as cluster \emph{g3}.}
  \label{fig:result_ngc_2516_dec}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_pm_filtered_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_parallax_filtered_ngc_2516.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2516/dec_hr_diagram_filtered_ngc_2516.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2516 DEC (filtered) characterization. Identified as cluster \emph{g3}.}
  \label{fig:result_ngc_2516_dec_filtered}
\end{figure}

@@

Figure \ref{fig:result_ngc_2516_clusterix} shows NGC 2516 characterized using Clusterix+TOPCAT tools.
Figure \ref{fig:result_ngc_2516_kmeans} shows eight clusters identified by K-Means.
In this case, cluster \emph{g3} is the one we are looking for.
Figures \ref{fig:result_ngc_2516_dec} and \ref{fig:result_ngc_2516_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
Again, the cluster of interest is the group \emph{g3}.
Although in general, groups between K-Means model and DEC models may not match.

@@

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & -4.6579 $\pm$ 0.0075 & 11.1517 $\pm$ 0.0075 & 2.4118 $\pm$ 0.0006 & 1727 \\
      Clusterix+TOPCAT & -4.652 $\pm$ 0.523 & 11.203 $\pm$ 0.454 & 2.409 $\pm$ 0.127 & 638 \\
      K-Means & -4.344 $\pm$ 0.14 & 9.507 $\pm$ 0.19 & 2.268 $\pm$ 0.01 & 1542 \\
      DEC & -4.426 $\pm$ 0.17 & 9.952 $\pm$ 0.20 & 2.436 $\pm$ 0.01 & 1532 \\
      \textbf{DEC (filtered)} & -4.502 $\pm$ 0.14 & 10.114 $\pm$ 0.17 & 2.392 $\pm$ 0.004 & 1072 \\
    \end{tabular}
    \caption{NGC 2516 results.}
    \label{tab:results_ngc_2516}
  \end{center}
\end{table}

@@

\begin{figure}[H]
  \vfil
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/pm_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/parallax_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/hr_diagram_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 Clusterix+TOPCAT characterization.}
  \label{fig:result_ngc_2632_clusterix}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_pm_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_parallax_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/kmeans_hr_diagram_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 K-Means characterization. Identified as cluster \emph{g1}.}
  \label{fig:result_ngc_2632_kmeans}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 DEC characterization. Identified as cluster \emph{g1}.}
  \label{fig:result_ngc_2632_dec}
  \centering
  \begin{subfigure}{0.9\textwidth}
    \centering
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_pm_filtered_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_parallax_filtered_ngc_2632.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.30\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/ngc_2632/dec_hr_diagram_filtered_ngc_2632.png}
    \end{subfigure}
  \end{subfigure}
  \caption{NGC 2632 DEC (filtered) characterization. Identified as cluster \emph{g1}.}
  \label{fig:result_ngc_2632_dec_filtered}
\end{figure}

@@

Figure \ref{fig:result_ngc_2632_clusterix} shows NGC 2632 characterized by the validation method.
Figure \ref{fig:result_ngc_2632_kmeans} shows ten clusters identified by K-Means.
Cluster \emph{g1} is the open cluster we are looking for.
Figures \ref{fig:result_ngc_2632_dec} and \ref{fig:result_ngc_2632_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
Open cluster NGC 2632 is labeled as group \emph{g1}.

@@

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & -36.047 $\pm$ 0.110 & -12.917 $\pm$ 0.066 & 5.371 $\pm$ 0.003 & - \\
      Clusterix+TOPCAT & -36.154 $\pm$ 1.001 & -12.909 $\pm$ 0.806 & 5.327 $\pm$ 0.187 & 371 \\
      K-Means & -26.352 $\pm$ 0.82 & -15.828 $\pm$ 0.76 & 5.394 $\pm$ 0.03 & 629 \\
      DEC & -20.012 $\pm$ 0.69 & -14.742 $\pm$ 0.58 & 4.686 $\pm$ 0.03 & 894 \\
      \textbf{DEC (filtered)} & -21.571 $\pm$ 0.74 & -14.234 $\pm$ 0.61 & 4.719 $\pm$ 0.03 & 714 \\
    \end{tabular}
    \caption{NGC 2632 results.}
    \label{tab:results_ngc_2632}
  \end{center}
\end{table}

@@

Figure \ref{fig:result_ngc_2682_clusterix} shows NGC 2682 characterized using Clusterix+TOPCAT tools.
Figure \ref{fig:result_ngc_2682_kmeans} shows ten clusters identified by K-Means.
K-Means labels NGC 2682 as group \emph{g4}.
Figures \ref{fig:result_ngc_2682_dec} and \ref{fig:result_ngc_2682_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
The cluster of interest is the group \emph{g2}.

@@

\begin{table}[h]
  \begin{center}
    \begin{tabular}{l|c|c|c|c}
      \textbf{Source / method} & \emph{$\mu_{\alpha}$ $(mas/yr^{-1})$} & \emph{$\mu_{\delta}$ $(mas/yr^{-1})$} & \emph{$\varpi$ $(mas)$} & \emph{\# stars} \\
      \hline
      \textbf{Simbad.u-strasbg.fr} & -10.9737 $\pm$ 0.0064 & -2.9396 $\pm$ 0.0063 & 1.1325 $\pm$ 0.0011 & 1194 \\
      Clusterix+TOPCAT & -10.970 $\pm$ 0.322 & -2.958 $\pm$ 0.327 & 1.142 $\pm$ 0.080 & 649 \\
      K-Means & -8.616 $\pm$ 0.15 & -3.710 $\pm$ 0.16 & 1.196 $\pm$ 0.01 & 1374 \\
      DEC & -8.926 $\pm$ 0.15 & -3.550 $\pm$ 0.15 & 1.144 $\pm$ 0.005 & 1238 \\
      \textbf{DEC (filtered)} & -9.619 $\pm$ 0.13 & -3.317 $\pm$ 0.13 & 1.140 $\pm$ 0.003 & 990 \\
    \end{tabular}
    \caption{NGC 2682 results.}
    \label{tab:results_ngc_2682}
  \end{center}
\end{table}

@@

Figure \ref{fig:result_melotte_25_clusterix} shows Melotte 25 characterized using Clusterix+TOPCAT tools.
Figure \ref{fig:result_melotte_25_kmeans} shows nine clusters identified by K-Means.
K-Means labels Melotte 25 as group \emph{g3}.
Figures \ref{fig:result_melotte_25_dec} and \ref{fig:result_melotte_25_dec_filtered}
show the groups found using the DEC model and the DEC model filtered, respectively.
Melotte 25 is labeled as group \emph{g5} by DEC model.

@@

We have applied our proposed method to clusters with different typologies:
